Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics and Predicts Long Timescale Behavior in Nanoscale Pores.

### 2019-12-04 Ben_Manuscripts/stochastic_transport/Draft.tex
Added databases for levy corrections

  \subsection{Mean First Passage Time}\label{method:mfpt}

  We used the mean first passage time (MFPT) in order to quantify how long it would take a 
  typical particle, parameterized with the above models, to travel across a membrane pore
  of length $L$. To measure this quantity, we simulated 50000 realizations of particle 
  trajectories, released at the pore entrance ($z=0$). In the case of uncorrelated hops, one
  can continuously draw from the hop length distribution until $z \geq L$ (or $-L$ for 
  the sake of computational efficiency). The length of time between the last time the 
  particle crossed $z=0$ and the end of the trajectory gives a single passage time. 
  When particle hops are correlated, as they are in all cases of this work, we cannot 
  continuously construct the particle trajectories. Rather, we must generate trajectories
  of length $n$ and measure the length of the sub-trajectory which traverses from $0$ 
  to $L$ without becoming negative.
  %BJC4: below is only necessary if you the mean of the distribution exists.
  %It is important to choose a value of $n$ sufficiently
  %large that all trajectories reach $L$, or else the mean first passage time will 
  %be underestimated.

  Our definition of the MFPT differs from more standard treatments. The setup of our 
  system is consistent with a one dimensional interval with two absorbing boundaries. In
  the absence of an external force and with a constant diffusion coefficient,
  $D$, the MFPT is simply:
  \begin{equation}
  T = \frac{z_0(L - z_0)}{2D}
  \end{equation}
  where $z_0$ is the initial position of particle trajectories.~\cite{gitterman_mean_2000}
  Of course that implies a conventional MFPT of 0 since particles overwhelmingly exit the
  pore at $z=0$. Therefore, we chose only to include trajectories in the MFPT calculation 
  which cross the boundary at $L$. We are unaware of an analytical form of the MFPT or 
  the distribution of passage times describing our system even in the simplest case of 
  Brownian increments. %BJC4: still digging for an analytical solution. Rich said the 
  %'marathon' runner problem might give a solution. It may be the same as the distribution
  % of runners crossing the finish line

  %BJC4: I'm not sure the best way to get the peak
  In the absence of an analytical equation for the MFPT, we instead report the location of
  the peak of the passage time distribution. This has the added advantage of treating
  cases where the passage time distribution diverges and no mean exists. While we will
  call the peak in the distribution the MFPT, it is important to recognize that this 
  value is more representative of a typical passage time rather than a mean.

  \begin{equation}
  F(t) = erfc\bigg(\frac{x - vt}{2\sqrt{D t}}\bigg)
  \end{equation}

>>>>

  \subsection{Solute Flux}\label{method:mfpt}

  We determined the rate at which solutes cross macroscopic-length pores based on the
  Hill relation:~\cite{hill_free_1989}
  \begin{equation}
  J = \frac{1}{MFPT}
  \end{equation}
  where $J$ is solute flux and MFPT refers to the mean first passage time. In the 
  context of our work, the MFPT describes the average length of time it takes a particle
  to move from the pore entrance to the pore exit.

  We generated particle trajectories, parameterized with the above models, in order
  to construct a distribution of first passage times across a membrane pore of length $L$.
  We simulated 50000 trajectory realizations, released at the pore entrance ($z=0$). In the case of 
  uncorrelated hops, one can continuously draw from the hop length distribution until 
  $z \geq L$ (or $-L$ for the sake of computational efficiency). The length of time 
  between the last time the particle crossed $z=0$ and the end of the trajectory gives a 
  single passage time. When particle hops are correlated, as they are in all cases of
  this work, we cannot continuously construct the particle trajectories. Rather, we must
  generate trajectories of length $n$ and measure the length of the sub-trajectory 
  which traverses from $0$ to $L$ without becoming negative.
  %BJC4: below is only necessary if you the mean of the distribution exists.
  %It is important to choose a value of $n$ sufficiently
  %large that all trajectories reach $L$, or else the mean first passage time will 
  %be underestimated.

  We calculated the expected value of analytical fits to the passage time distributions
  in order to determine the MFPT for a given solute and pore length. One should not 
  use the mean of the empirical passage time distribution because it is highly likely
  that the true MFPT will be underestimated unless 100\% of a very large number of 
  trajectories reach $L$. If a trajectory does not reach $L$ within $n$ steps, it is
  possible that a very long passage time has been excluded from the distribution.

  %BJC4: can probably delete this paragraph
%  We are unaware of an analytical passage time distribution which describes our 
%  geometry. The setup of our system is consistent with a one dimensional interval 
%  with two absorbing boundaries. In the absence of an external force and with a constant
%  diffusion coefficient, $D$, the MFPT is simply:
%  \begin{equation}
%  T = \frac{z_0(L - z_0)}{2D}
%  \end{equation}
%  where $z_0$ is the initial position of particle trajectories.~\cite{gitterman_mean_2000}
%  Of course that implies a conventional MFPT of 0 since particles overwhelmingly exit the
%  pore at $z=0$. Therefore, we chose only to include trajectories in the MFPT calculation 
%  which cross the boundary at $L$. We are unaware of an analytical form of the MFPT or 
%  the distribution of passage times describing our system even in the simplest case of 
%  Brownian increments. %BJC4: still digging for an analytical solution. Rich said the 
%  %'marathon' runner problem might give a solution. It may be the same as the distribution
%  % of runners crossing the finish line

  To come up with an analytical equation describing the passage time distributions, 
  one can frame the problem as a pulse of particles at the pore inlet ($z=0$) that 
  spreads out over time as it approaches $L$.

%  This pulse model effectively describes
%  the time-dependent probability distribution of particles which eventually reaches 
%  $L$. The diffusion equation describing this situation is:
%  \begin{equation}
%  \frac{\partial c}{\partial t} + v\frac{\partial c}{\partial x} = D\frac{\partial^2 c}{\partial x^2}
%  \end{equation}
%  where $v$ is some average velocity describing the location of the center of the
%  distribution and $D$ is a diffusion constant. 
%  One can show that the normalized
%  concentration profile is:
  First, consider a pulse spreading out with a fixed mean. It is well known that the 
  time-dependent probability distribution of particle positions is:
  \begin{equation}
  p(x, t) = \frac{1}{\sqrt{4 \pi D t}}exp\bigg(\frac{-x^2}{4Dt}\bigg)
  \end{equation}
  We can make the substitution $x = x - vt$ in order to linearly shift the mean as a
  function of time:
  \begin{equation}
  p(x, t) = \frac{1}{\sqrt{4 \pi D t}}exp\bigg(\frac{-(x - vt)^2}{4Dt}\bigg)
  \end{equation}
  One can track the number of the fraction of particles, $F$, that have crossed the pore 
  boundary by integrating:
  \begin{equation}
  F = \int_x^\infty p~ dx = erfc\bigg(\frac{x - vt}{2\sqrt{D t}}\bigg)
  \end{equation}
  and finally take the derivative to arrive at the first passage time distribution:
  \begin{equation}
  P = -\frac{1}{\sqrt{\pi}}e^{-(x - vt)^2 / (4Dt)}\bigg(-\frac{D(x - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \end{equation}

  %BJC4: we have better ways of getting peaks now
%  In the absence of an analytical equation for the MFPT, we instead report the location of
%  the peak of the passage time distribution. This has the added advantage of treating
%  cases where the passage time distribution diverges and no mean exists. While we will
%  call the peak in the distribution the MFPT, it is important to recognize that this 
%  value is more representative of a typical passage time rather than a mean.
